How to \emph{not} write Virtual Machines for Dynamic Languages

### 2007-07-24 talk/dyla2007/talk.tex
somewhat uninspired pypy section

\begin{frame}
  \frametitle{}
  \begin{itemize}
  \item
  \end{itemize}
  \pause
  \begin{block}{
    Python Case}
    \begin{itemize}
    \item
    \end{itemize}
  \end{block}
\end{frame}

>>>>

\begin{frame}
  \frametitle{PyPy's Approach to VM Construction}
  \begin{itemize}
  \item
    Goal: achieve flexibility, simplicity and performance all at one
  \item
    Approach: auto-generate VMs from high-level descriptions of the language
  \item
    ... using meta-programming techniques and aspects (in the original sense)
  \item
    high-level description of the language done by writing an interpreter in
    RPython
  \end{itemize}
  \pause
  \begin{block}{
    What is RPython}
    \begin{itemize}
    \item
      RPython is a subset of Python
    \item
      subset chosen in such a way that type-inference can be performed
    \item
      still a high-level language (unlike SLang or Prescheme)
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Auto-generating VMs}
  \begin{itemize}
  \item
    interpreters written in RPython can be translated to various targets:
    C/Posix, CLR, JVM
  \item
    VMs generated by a flexible translation toolchain (really a compiler)
  \end{itemize}
  \begin{block}{
    Translation Aspects}
    \begin{itemize}
    \item
      the translation toolchain weaves in various aspect into the generated VM
    \item
      examples: GC strategy, threading model
    \item
      early design decisions not necessary
    \item
      non-trivial translation aspect: auto-generating a dynamic compiler from
      the interpreter % XXX maybe we really want to write more about this
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Good Points of the Approach}
  \begin{itemize}
  \item
    dynamic languages can implemented in a high level language
  \item
    doing this separates high-level from low-level concerns
  \item
    generating various VMs from a single source interpreter, which is easy to
    write
  \item
    lots of flexibility at all levels:
    \begin{itemize}
    \item
      when writing the interpreter
    \item
      also possible to adapt the translation toolchain if necessary
    \item
      this allows breaking of abstraction barriers, when necessary (e.g. tagged
      integers)
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Open Issues / Drawbacks / Further Work}
  \begin{itemize}
  \item
    writing the translation toolchain in the first place takes lots of effort
    (but it can be reused)
  \item
    writing a good GC is still necessary. but: can we maybe reuse existing good
    GCs (e.g. from the Jikes RVM)?
  \item
    many abstraction layers
  \item
    dynamic compiler generation seems to work, but needs to be improved
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Conclusion / Meta-Points}
  \begin{itemize}
  \item
    high-level languages are suitable to implement dynamic languages
  \item
    doing so has many benefits
  \item
    VM's shouldn't be written by hand
  \item
    PyPy's concrete approach is not so important
  \item
    diversity is good
  \item
    let's write more meta-programming tool-chains!
  \end{itemize}
\end{frame}
